Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network

### 2017-11-01 www-2018/source/paper.tex
Final revisions.

Under the above assumptions,
we show how to evaluate the influence of network structure
on attack-tolerance.
We next present a structured multipath fault tolerance (SMFT) to extend
standard fault tolerance techniques to the case of adversarial faults in
networks
\cite{avizienis_basic_2004, von_neumann_probabilistic_1956}.
The SMFT scheme requires the existence of a
concurrent multipath routing (CMR) algorithm
\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013},
to takes advantage of the independence of faults
along {\em independent paths}.
We also present a novel CMR algorithm for the butterfly network topology.
The butterfly topology is popular in parallel processing
\cite{kshemkalyani_distributed_2008} and
peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
applications, due to its regular structure, low degree, and high connectivity.

It is important to note that the butterfly is a highly structured and constrained
network topology,
very different form those found in social networks and other
self-organized networks.
The reader may wonder whether it is realistic or useful to assume such control over
the network structure.
We have already seen that whenever a single point of failure exists in the network,
there is a potential for an attacker to exploit it through coercion.
So, {\em attack-tolerance cannot be achieved without the ability to influence
network structure}.
Luckily, there are scenarios in which network topology can be dictated.
Examples include overlay networks
\cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
In general,
{\em when the need for attack-tolerance is high enough to warrant investment
in infrastructure, networks can be engineered and maintained as infrastructure}.
It is also worth noting that attack-tolerant networks may be sub-components of
larger, less-constrained systems.
For example, a single server might be replaced by a distributed network of servers,
each with different ownership, physical location, and legal jurisdiction,
without placing any unrealistic constraints on the clients connecting to
those servers.

Our main contributions are:
\begin{itemize}
\item{
We propose a novel structured multipath fault tolerance (SMFT) scheme
for extending standard fault tolerance techniques to
{\em adversarial} faults in {\em complex networks}.
We do so by modeling redundant {\em independent paths}
with {\em bounded trust transitivity} as a single virtual channel,
and show that the probability of detecting adversarial faults
approaches 1 exponentially with the number of paths;
}
\item{We prove that the number of $h$-internally vertex disjoint
paths between two nodes in a wrap-around butterfly network
increases exponentially with $h$,
and thus that the number of independent paths in a network
increases exponentially with the level of trust transitivity;
}
\item{We present a scalable and efficient concurrent
multipath routing (CMR) algorithm on the butterfly network topology,
which can be combined with SMFT to achieve a high level of attack tolerance.}
\end{itemize}

This paper is organized as follows.
Section \ref{sec-related} reviews background and related work.
Section \ref{sec-ft} describes adversarial fault tolerance on
structured networks.
Section \ref{sec-butterfly} gives background on the butterfly network topology.
Section \ref{sec-bf-route} presents our concurrent multipath routing
algorithm for the butterfly.
Section \ref{sec-discussion} discusses the results.
And Section \ref{sec-conclusion} concludes.

@@

Standard fault tolerance methods use redundancy to detect and correct
statistically indpendent faults.
In complex networks however, faults can be correlated when,
for example, two messages pass through the same faulty node.
For now, let us assume our sender (Alice) and receiver (Bob) are connected
by $\delta$ direct channels, with indpendent errors.
We will return to the question of constructing these channels in subsequent
sections.
For now, we concern ourselves with the question:
given that the network provides $\delta$ redundant channels between
Alice and Bob,
what is the probability that an adversary (Mal) causes an undetectable
error after causing faults in a fixed number of channels?

@@

So far, we have assumed that Alice and Bob have access to some number
$\delta$ of channels with statistically indpendent faults.
However, in real communication architectures, direct links between all
pairs of individuals not possible and messages must be routed through
a number of intermediate nodes.
In an adversarial setting,
the existence of intermediate nodes introduces two problems:
1. intermediate nodes may be compromised by the adversary and
2. faults on paths are no longer statistically independent:
two paths may pass through the same compromised node.
We show how to how to resolve these problems using a combination of network
structure and bounded trust transitivity.

@@

In order to implement structured multipath fault tolerance,
we need a strucutred network topology with high effective redundancy.
In this paper, we apply SMFT to the butterfly network topology
\cite{kshemkalyani_distributed_2008}.
The structure of the butterfly network is highly constrained,
making it most suitable for applications where portions of the
network structure can be designed or dictated.
Examples of such networks include:
overlay networks \cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
However, when attack-tolerance is desired,
it will always require control over network structure in order to
eliminate single points of failure.
The regular structure of the butterfly is not a limitation of our approach,
but rather a reflection of the inherent difficulty of attack-tolerance.
To address the case when the network cannot be fully controlled,
we show how partially rewiring a snapshot of the internet's router
network can greatly increase it's effective redundancy and
attack-tolerance properties.

@@

Even when a butterfly topology cannot be implemented perfectly,
it can still increase the attack tolerance properties of a network.
Here, we simulate targeted attacks against a snapshot of the internet's
router network on January 2, 2000
\cite{leskovec_graphs_2005}.
However, we equate a subset of the nodes with the nodes of a 9-dimensional
butterfly network and rewire a fraction of the edges to match the butterfly
topology.
At each step of the simulation, betweenness centrality is recalculated and the
most central node is removed.

Our simulations show improved resistance to fragmentation and higher
effective redundancy when even a fraction of edges have been rewired to
match the butterfly topology.
While the original router network fragments when about 1\% of the nodes
have been removed (Figure \ref{fig:diameter}),
this number increases to 2\%
with only 10% of the edges rewired to match the butterfly.
With 90\% of the edges rewired,
the network remains unfragmented beyond the failure of the 8\% most
central nodes.
The effective redundancy for varius values of trust transitivity $h$
are shown in Figure \ref{fig:mincut}.
The effective redundancy is calculated by collapsing nodes and edges within
$h$ hops of source and destination into single nodes and finding the mincut
between them,
averaging over 150 source-destination pairs.
For all but $h=0$,
the original network has higher redundancy for a small fraction of
node failures,
while the rewired network has higher redundancy beyond a crossover point.
We interpret these results to suggest that even when a small number of
highly central nodes have been removed from a heavy-tailed network,
most nodes are still able to take advantage of the remaining hubs.
As larger hubs continue to be removed, the connectivity of the network
decreases until the crossover point,
at which point the rewired network offers higher effective redundancy.

\begin{figure}
\centerline{\includegraphics[width=3in,height=2.25in]{fig-diameter}}
\caption{
Simulation of targeted attacks against a snapshot of the internet's router
network with a fraction of the edges rewired into a partial butterfly configuration.
The original network fragments at about 1\%.
With 90\% of the edges rewired,
the network remains unfragmented beyond the failure of the 8\% most
central nodes.
}
\label{fig:diameter}
\end{figure}

>>>>

Under the above assumptions,
we show how to evaluate the influence of network structure
on attack-tolerance.
We next present a structured multipath fault tolerance (SMFT) scheme to extend
standard fault tolerance techniques to the case of adversarial faults in
networks
\cite{avizienis_basic_2004, von_neumann_probabilistic_1956}.
The SMFT scheme requires the existence of a
concurrent multipath routing (CMR) algorithm
\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013},
to takes advantage of the independence of faults
along {\em independent paths}.
We also present a novel CMR algorithm for the butterfly network topology.
The butterfly topology is popular in parallel processing
\cite{kshemkalyani_distributed_2008} and
peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
applications, due to its regular structure, low degree, and high connectivity.

It is important to note that the butterfly is a highly structured and constrained
network topology,
very different form those found in social networks and other
self-organized networks.
The reader may wonder whether it is realistic or useful to assume such control over
the network structure.
We have already seen that whenever a single point of failure exists in the network,
there is a potential for an attacker to exploit it through coercion.
So we claim that {\em attack-tolerance cannot be achieved without the ability to influence
network structure}.
Luckily, there are scenarios in which network topology can be dictated.
Examples include overlay networks
\cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
In general,
{\em when the need for attack-tolerance is high enough to warrant investment
in infrastructure, networks can be engineered and maintained as infrastructure}.
It is also worth noting that attack-tolerant networks may be sub-components of
larger, less-constrained systems.
For example, a single server might be replaced by a distributed network of servers,
each with different ownership, physical location, and legal jurisdiction,
without placing any unrealistic constraints on the clients connecting to
those servers.

Our main contributions are:
\begin{itemize}
\item{
We propose a novel structured multipath fault tolerance (SMFT) scheme
for extending standard fault tolerance techniques to
{\em adversarial} faults in {\em complex networks}.
We do so by showing that the probability of detecting adversarial faults
approaches 1 exponentially with the number of {\em independent paths},
and that with $h$-degree {\em bounded trust transitivity},
all $h$-internally vertex disjoint paths are independent.
}
\item{We prove that the number of $h$-internally vertex disjoint
paths between two nodes in a wrap-around butterfly network
is at least $2^h$,
and present a scalable and efficient concurrent multipath routing (CMR) algorithm
to find these paths,
which can be combined with SMFT to achieve a high level of attack-tolerance.
}
\item{We show that rewiring a snapshot of the internet's router network with
edges from a butterfly network allows it to tolerate a higher number of failures
without fragmenting, and increases the effective redundancy in the presence
of a large number of adversarial faults.}
\end{itemize}

This paper is organized as follows.
Section \ref{sec-related} reviews background and related work.
Section \ref{sec-ft} describes adversarial fault tolerance on
structured networks.
Section \ref{sec-butterfly} gives background on the butterfly network topology.
Section \ref{sec-bf-route} presents our concurrent multipath routing
algorithm for the butterfly network.
Section \ref{sec-discussion} discusses the results.
And Section \ref{sec-conclusion} concludes.

@@

Standard fault tolerance methods use redundancy to detect and correct
statistically independent faults.
In complex networks however, faults can be correlated when,
for example, two messages pass through the same faulty node.
For now, let us assume our sender (Alice) and receiver (Bob) are connected
by $\delta$ direct channels, with indpendent errors.
We will return to the question of constructing these channels in subsequent
sections.
For now, we concern ourselves with the question:
given that the network provides $\delta$ redundant channels between
Alice and Bob,
what is the probability that an adversary (Mal) causes an undetectable
error after causing faults in a fixed number of channels?

@@

So far, we have assumed that Alice and Bob have access to some number
$\delta$ of channels with statistically indpendent faults.
However, in real communication architectures, direct links between all
pairs of individuals are not possible and messages must be routed through
a number of intermediate nodes.
In an adversarial setting,
the existence of intermediate nodes introduces two problems:
1. intermediate nodes may be compromised by the adversary and
2. faults on paths are no longer statistically independent:
two paths may pass through the same compromised node.
We show how to how to resolve these problems using a combination of network
structure and bounded trust transitivity.

@@

In order to implement structured multipath fault tolerance,
we need a strucutred network topology with high effective redundancy.
In this paper, we apply SMFT to the butterfly network topology
\cite{kshemkalyani_distributed_2008}.
The structure of the butterfly network is highly constrained,
making it most suitable for applications where portions of the
network structure can be designed or dictated.
Examples of such networks include:
overlay networks \cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
More flexible architectures may be possible,
but attack-tolerance will always require some level of influence over
network structure in order to limit single points of failure.
To address the case when the network cannot be fully controlled,
we show how partially rewiring a snapshot of the internet's router
network can greatly increase it's effective redundancy and
attack-tolerance properties.

@@

Even when a butterfly topology cannot be implemented perfectly,
it can still increase the attack tolerance properties of a network.
Here, we simulate targeted attacks against a snapshot of the internet's
router network on January 2, 2000
\cite{leskovec_graphs_2005}, having 6493 nodes and 13914 edges.
At each step of the simulation, betweenness centrality is recalculated and the
most central node is removed.
We also simulate attack on several rewired networks.
We 1. generate edges corresponding to a 9-dimensional butterfly network between
the 4608 highest-degree router nodes,
2. choose a fraction $f$ of those edges at random,
3. add those edges to the router network, and
4. remove an equal number of the original edges at random.

Our simulations show improved resistance to fragmentation and higher
effective redundancy when even a fraction of edges have been rewired to
match the butterfly topology.
While the original router network fragments when about 1\% of the nodes
have been removed (Figure \ref{fig:diameter}),
this number increases to 2\%
with only 10\% of the butterfly edges present.
With 90\% of the butterfly edges present
the network remains unfragmented beyond the failure of the 8\% most
central nodes.
The effective redundancy for varius values of trust transitivity $h$
are shown in Figure \ref{fig:mincut}.
The effective redundancy is calculated by collapsing nodes and edges within
$h$ hops of source and destination into single nodes and finding the mincut
between them,
averaging over 150 source-destination pairs.
For $h=0$, the rewired version has strictly higher redundancy.
For $h>0$,
the original network has higher redundancy when the number of failures is small,
while the rewired network has higher redundancy beyond a crossover point.
We interpret these results to suggest that even when a small number of
highly central nodes have been removed from a heavy-tailed network,
most nodes are still able to take advantage of the remaining hubs.
As larger hubs continue to be removed, the connectivity of the network
decreases until the crossover point,
at which point the rewired network offers higher effective redundancy.

\begin{figure}
\centerline{\includegraphics[width=3in,height=2.25in]{fig-diameter}}
\caption{
Simulation of targeted attacks against a snapshot of the internet's router
network with a fraction of the edges rewired into a partial butterfly configuration.
The original network fragments at about 1\%.
With 90\% of the butterfly edges present,
the network remains unfragmented beyond the failure of the 8\% most
central nodes.
}
\label{fig:diameter}
\end{figure}
